
{\color{red} \bf Warning:}
The findings and conclusions in this article have not been
formally disseminated by the U.S. Department of Health \& Human Services,
U.S. Food and Drug Administration, nor Oak Ridge Institute for Science and
Education. They should not be construed to represent any determination or
policy of University, Agency, Administration and National Laboratory.

This document is written to explain the main
functions of \pkg{pbdZMQ}~\citep{Chen2015}, version 0.1-0.
Every effort will be made to ensure future versions are consistent with
these instructions, but features in later versions may not be explained
in this document.

Information about the functionality of this package,
and any changes in future versions can be found on website:
``Programming with Big Data in R'' at
\url{http://r-pbd.org/}~\citep{pbdR2012}.




\section[Introduction]{Introduction}
\label{sec:introduction}
\addcontentsline{toc}{section}{\thesection. Introduction}


ZeroMQ (ØMQ) at \url{http://www.zeromq.org/} is a library for high-performance
asynchronous messaging in scalable distributed applications. It provides
APIs in several messaging patterns that essentially the basic way to
form connections between devices including computers, mobile devices, and
of course supercomputers. The APIs also simplify the complex calls to sockets
and reduce the burden of knowing network communications.
Several popular programming languages have also provided biding to utilize
and take advantage from those APIs.

In \pkg{pbdZMQ}, those ZeroMQ APIs are in detail manners wrapped in \proglang{R}
via lower level \proglang{C} language. \pkg{pbdZMQ} mainly implements
on a few ZeroMQ patterns including
\begin{itemize}
\item
request-reply, in particular, one client and a server, and
\item
push-pull, in particular, one client and a set of servers.
\end{itemize}
These patterns are useful communication frameworks designed in
\pkg{pbdCS}~\citep{Schmidt2015} that combines two different messaging libraries
ZeroMQ and MPI and utilizes their advantages in
\begin{itemize}
\item
interaction of user interface by ZeroMQ or \pkg{pbdZMQ}, and
\item
computation of statistical programming by MPI or \pkg{pbdMPI}~\citep{Chen2012}.
\end{itemize}
