\section[Solaris 10]{Solaris 10}
\label{sec:solaris_10}
\addcontentsline{toc}{section}{\thesection. Solaris 10}

Solaris 10 with external ZeroMQ 4.0.7 library and GNU tools (OpenCSW)
has be tested for \pkg{pbdZMQ} successfully. The \code{INSTALL.solaris}
has the steps which basically are
\begin{enumerate}
\item install GNU tools from OpenCSW using \code{pkgutil},
\item set path to GNU tools,
\item install external ZeroMQ 4.0.7 library using GNU tools, and
\item install \pkg{pbdZMQ} linked with external ZeroMQ.
\end{enumerate}
Next, I write down some considerations as a note.


\subsection[Have a Solaris 10?]{Have a Solaris 10?}
\label{sec:have_solaris_10}
\addcontentsline{toc}{subsection}{\thesubsection. Have a Solaris 10?}
It is easy and fun.
\begin{enumerate}
\item
Install an Oracle VM VirtualBox which is available at
\url{https://www.virtualbox.org/}.
\item
Download a Solaris template from \\
\url{http://www.oracle.com/technetwork/server-storage/vm/solaris-098101.html}.
\item
Import the template to the VM.
\item
Boot and setup root password, internet, share drives, etc.
\item
Done.
\end{enumerate}


\subsection{OpenCSW}{OpenCSW}
\label{sec:opencsw}
\addcontentsline{toc}{subsection}{\thesubsection. OpenCSW}
The OpenCSW (\url{www.opencsw.org}) provides binary packages of GNU tools
for Solaris 10. Ideally, gmake, gcc, g++, autoconf, automake, gtar are
required which all availalbe from the website. For example, the gcc is in
"gcc4core" package which can be installed with a few \code{pkgutil}
commands by
\begin{Code}
> pkgadd -d http://get.opencsw.org/now
> /opt/csw/bin/pkgutil -U
> /opt/csw/bin/pkgutil -y -i gcc4core 
> /usr/sbin/pkgchk -L CSWgcc4core # list files
\end{Code}
Typically, the binary packages are installed under
\code{/opt/csw/}.

Make sure those GNU tools are also in the path and export it to
the environment. For example, I have
\begin{Code}
PATH=/opt/csw/gnu:/opt/csw/bin:/opt/csw/i386-pc-solaris2.10/bin:
/usr/sbin:/usr/bin:/usr/openwin/bin:/usr/ucb
export PATH
\end{Code}
in a file name \code{set_path.sh}. Then, I set it to the environment by
\begin{Code}
> . ./set_path.sh
\end{Code}
I also have a symbolic link "make" to "gmake" to avoid configuration problems.
If not, use \code{ln -s} to make one.


\subsection{Compile ZeroMQ}{Compile ZeroMQ}
\label{sec:compile_zeromq}
\addcontentsline{toc}{subsection}{\thesubsection. Compile ZeroMQ}
The OpenCSW also has a binary package for ZeroMQ, named \pkg{libzmq1} or
\pkg{libzmq1\_dev}.
The package is a little bit out of date (version 2.2.0 built in 2012).
This old version is not comparable with current \pkg{pbdZMQ}.
The solution is to compile and install an external ZeroMQ 4.0.7
which available from \url{http://download.zeromq.org}.
This version 4.0.7 is the latest stable ZeroMQ version.
The newer versions of ZeroMQ, such as >= 4.1.0, may have more bugs
for Solaris 10.

I install the ZeroMQ library to \code{/work-my/local/zmq} with
the following commands.
\begin{Code}
> gtar zxvf zeromq-4.0.7.tar.gz
> cd zeromq-4.0.7
> ./configure --prefix=/work-my/local/zmq 
> make
> make install
\end{Code}
There may have some warnings, but they are ok. The default installs
the share library which can be linked by \pkg{pbdZMQ}.


\subsection{Install pbdZMQ}{Install pbdZMQ}
\label{sec:install_pbdzmq}
\addcontentsline{toc}{subsection}{\thesubsection. Install pbdZMQ}
With the OpenCSW and external ZeroMQ 4.0.7 installed, I can install
\pkg{pbdZMQ} simply via
\begin{Code}
> R CMD INSTALL pbdZMQ_0.2-0.tar.gz \
  --configure-vars="ZMQ_INCLUDE='-I/work-my/local/zmq/include' \
                    ZMQ_LDFLAGS='-L/work-my/local/zmq/lib -lzmq'"
\end{Code}
Then, the test can be done by
\begin{Code}
> Rscript -e "demo(hwserver,'pbdZMQ',ask=F,echo=F)" &
> Rscript -e "demo(hwclient,'pbdZMQ',ask=F,echo=F)"
\end{Code}
when the \pkg{pbdZMQ} installation is done correctly.

